Local Spectral Clustering of Density Upper Level Sets

### 2019-06-19 neurips/local_spectral_density_clustering_arxiv.tex
more random edits

In this paper, we study the problem of clustering: splitting a given data set
into groups that satisfy some notion of within-group similarity and
between-group difference.  We focus on spectral clustering, a family of powerful
nonparametric clustering algorithms.  Generally speaking, a spectral algorithm
first constructs a geometric graph $G$, where vertices correspond to samples,
and edges correspond to proximities between samples. It then learns a feature
embedding based on the Laplacian of $G$, and applies a simple clustering
technique (like k-means clustering) in the embedded feature space.

When applied to geometric graphs built from a large number of samples,
global spectral clustering methods can be computationally cumbersome and   
insensitive to the local geometry of the underlying distribution
\citep{leskovec2010,mahoney2012}.  This has led to increased interest in
local spectral clustering algorithms, which leverage locally-biased spectra
computed using random walks around some user-specified seed node.  A popular 
local clustering algorithm is Personalized PageRank (PPR), first introduced by 
\citet{haveliwala2003}, then further developed by
\citep{spielman2011,spielman2014,andersen2006,mahoney2012,zhu2013},
among others.

Local spectral clustering techniques have been practically very successful
\citep{leskovec2010,andersen2012,gleich2012,mahoney2012,wu2012}, leading 
many authors to develop supporting theory
\citep{spielman2013,andersen2009,gharan2012,zhu2013} that gives worst-case
guarantees on traditional graph-theoretic notions of cluster quality (such as
conductance).  In this paper, we adopt a more traditional statistical viewpoint,
and examine what the output of local clustering on a data set reveals about the
underlying density $f$.  In particular, we examine the ability of PPR to recover
\emph{density clusters} of $f$, defined as the connected components of
the upper level set $\{x \in \Rd : f(x) \geq \lambda\}$ for some $\lambda > 0$
(a central object of interest in the statistical clustering literature, dating
back to \citet{hartigan1981}).

@@

\paragraph{Summary of results.} A summary of our results (and outline for this
paper) is as follows.

@@

More broadly, density clustering and level set estimation is a well-studied
problem. \citet{polonik1995, rigollet2009} study density clustering under the 
symmetric set difference metric, \citet{tsybakov1997, singh2009} describe
minimax optimal level set estimators under Hausdorff loss and
\citet{hartigan1981, chaudhuri2010} consider consistent estimation of the
cluster tree, to note but a few works. Our goal is not to improve on these
results, nor to offer a better algorithm for level set estimation; indeed, seen as
a density clustering algorithm, PPR has none of the optimality guarantees 
found in the aforementioned works. Instead, our motivation is to start with a 
widely-used local spectral method, PPR, and to better understand and
characterize the distinctions between those density clusters which are
well-conditioned for PPR, and those which are not.

@@

\begin{theorem}
  \label{thm: misclassification_rate}
  Fix $\lambda > 0$, let \smash{$\Cset \in \Cbb_f(\lambda)$} be a
  $\kappa$-well-conditioned density cluster (with respect to some $\theta$), and
  additionally assume $f$ satisfies \ref{asmp: valid_region}. If Algorithm
  \ref{alg: ppr} is well-initialized, there exists a universal constant $c >
  0$ such that with probability tending to 1 as $n \to \infty$,  
  \begin{equation}
    \label{eqn: misclassification_rate_ub}
    \Delta(\Csig[\Xbf], \Cest) \leq c \kappa(\Cset)
    \frac{\Lambda_{\sigma}}{\lambda_{\min}}. 
  \end{equation}
\end{theorem}

@@

Neither \eqref{eqn: graph_symmetric_set_difference} nor \eqref{eqn:
  misclassification_rate_ub} imply consistent density cluster estimation in the 
sense of \eqref{eqn: consistent_density_cluster_recovery}. This notion of
consistency requires a uniform bound over $\pbf$: for all \smash{$\Cset'
  \in \Cbb_f(\lambda), \Cset' \neq \Cset$}, and each $u \in \Cset, w \in
\Cset'$,  
\begin{equation}
\label{eqn: ppr_gap}
\frac{p_{w}}{\Dbf_{ww}} \leq \frac{1}{40\vol_0} < \frac{1}{11\vol_0} \leq
\frac{p_u}{\Dbf_{uu}}, 
\end{equation}
so that any sweep cut $S_{\beta}$ for $\beta \vol_0 \in [1/40,1/11]$ (i.e., any
sweep cut considered by Algorithm \ref{alg: ppr}) will fulfill both conditions
laid out in \eqref{eqn: consistent_density_cluster_recovery}. In Theorem
\ref{thm: consistent_recovery_of_density_clusters}, we show that a sufficiently
small upper bound on $\kappa(\Cset)$ ensures such a gap exists with probability
1 as $n \to \infty$, and hence guarantees \smash{$\Cest$} will be a consistent  
estimator. As was the case before, we wish to preclude arbitrarily low degree
vertices, this time for points $x \in \Cset'[\Xbf]$. 
\begin{enumerate}[label=(A\arabic*)]
  \setcounter{enumi}{6}
\item 
  \label{asmp: C'_bounded_density}
  \emph{Bounded density in other clusters:} Letting $\sigma,\lambda_{\sigma}$ be 
  as in \ref{asmp: bounded_density}, for each $\Cset' \in \Cbb_f(\lambda)$ and
  for all $x \in \Csig'$, $\lambda_{\sigma} \leq f(x)$. 
\end{enumerate}

@@

We generalize some notation from the main text. For an undirected graph $G = (V,E)$, the lazy random walk over $G$ is the Markov chain with transition probabilities given by $\Wbf := \frac{\Ibf{} + \Dbf^{-1} \Abf}{2}$ and stationary distribution $\pi$.  Denote the $m$-step probability distribution of this random walk originating from a particular $v \in V$ as $q^{(m)}: V \times V \to [0,1]$, $q^{(m)}(v,u) = e_v \Wbf^m e_u$. An important intermediate quantity used to bound the relative pointwise mixing time $\tau_{\infty}(G)$ is the total variation distance between the distributions $q_v^{(m)} := q^{(m)}(v,\cdot)$ and $\pi$,
\begin{equation*}
\norm{q_v^{(m)} - \pi}_{TV} = \sum_{u \in V} \abs{q_v^{(m)}(u) - \pi(u)}
\end{equation*}
We recall the \emph{cut} and \emph{volume} functionals over a graph, and introduce the \emph{degree} functional as well. For $u \in V$, $S \subseteq V$, 
\begin{equation*}
\cut(S;G) = \sum_{u \in S} \sum_{v \in S^c} \1((u,v) \in E), \quad \deg(u;G) = \sum_{v \in V} \1((u,v) \in E), \quad \vol(S;G) = \sum_{u \in S} \deg(u;G)
\end{equation*}

@@

Proposition \ref{prop: tv_mixing_time} relates these geometric quantities to the TV distance between $q_v^{(t)}$ and stationary distribution $\pi$.
\begin{proposition}
	\label{prop: tv_mixing_time}
	For any $v \in V$, and any $0 < a \leq 1$,
	\begin{equation*}
	\norm{q_v^{(t + 3)} - \pi}_{TV} \leq \set{as(G), \frac{1}{8} + \frac{1}{20} + \frac{1}{2 \min_{u \in V} \deg(u;G)}} + \left(\frac{1}{1 - 2as(G)} + \frac{1}{2 as(G)}\right) \left(1 - \frac{\Phi^2(G)}{2}\right)^t
	\end{equation*}
\end{proposition}

@@

\begin{lemma}
	\label{lem: tv_to_uniform_distance}
	Let $\norm{q_v^{(t)} - \pi}_{TV} \leq \frac{1}{14} \max \set{1, \frac{1}{s(G)}}$. Then,
	\begin{equation*}
	d_{\mathrm{unif}}(q_v^{(t + 3)},\pi) \leq \frac{1}{4}
	\end{equation*}
\end{lemma}
\begin{proof}
	Fix $u \in V$ and let $m \geq t + 1$ be arbitrary. The stationarity of $\pi$ gives
	\begin{align}
	\frac{\pi(u) - q_v^{m}(u)}{\pi(u)} & = \sum_{y \in V} \left(\pi(y) - q^{(m-1)}(v,y)\right) \left(\frac{q^{(1)}(y,u) - \pi(u)}{\pi(u)}\right) \nonumber \\
	& \overset{\text{(i)}}{=} \sum_{y \neq u} \left(\pi(y) - q^{(m-1)}(v,y)\right) \left(\frac{q^{(1)}(y,u) - \pi(u)}{\pi(u)}\right) + \frac{\pi(u) - q^{(m - 1)}(v,u)}{\pi(u)} \left(\frac{1}{2} - \pi(u)\right) \nonumber \\
	& \leq \sum_{y \neq u} \left(\pi(y) - q^{(m-1)}(v,y)\right) \left(\frac{q^{(1)}(y,u) - \pi(u)}{\pi(u)}\right) + \frac{\pi(u) - q^{(m - 1)}(v,u)}{2 \pi(u)} \label{eqn: tv_to_uniform_distance_1}
	\end{align}
	where $\text{(i)}$ follows from $q^{(1)}(u,u) = \frac{1}{2}$.

	Then 
	\begin{align}
	\sum_{y \neq u} \left(\pi(y) - q^{(m-1)}(v,y)\right) \left(\frac{q^{(1)}(y,u) - \pi(u)}{\pi(u)}\right) & \leq \norm{q_v^{(m-1)} - \pi}_{TV} \max_{y \neq u} \abs{\frac{q^{(1)}(y,u) - \pi(u)}{\pi(u)}} \nonumber \\
	& \leq \norm{q_v^{(m-1)} - \pi}_{TV} \max \set{1, \max_{y \neq u}\set{\frac{q^{(1)}(y,u)}{\pi(u)}}} \nonumber \\
	& \leq \norm{q_v^{(m-1)} - \pi}_{TV} \max \set{1, \frac{1}{s(G)}} \label{eqn: tv_to_uniform_distance_3}
	\end{align}
	since for $y \neq u$, $q^{(1)}(y,u) \leq 1/\left(2 \min_{u \in V} \deg(u; G)\right)$. As $m - 1 \geq t$, it is well known \cite{lovasz1990} that the laziness of the random walk guarantees $\norm{q_v^{(m - 1)} - \pi}_{TV} \leq \norm{q_v^{(t)} - \pi}_{TV}$, and therefore by \eqref{eqn: tv_to_uniform_distance_3} and hypothesis,
	\begin{equation*}
	\sum_{y \neq u} \left(\pi(y) - q^{(m-1)}(v,y)\right) \left(\frac{q^{(1)}(y,u) - \pi(u)}{\pi(u)}\right) \leq \frac{1}{14}.
	\end{equation*}

@@

\begin{proposition}
	\label{prop: pointwise_mixing_time}
	Assume $\min_{u \in v} \deg(u; G) \geq 10$. Then,
	\begin{equation*}
	\tau_{\infty}(G) \leq \frac{2}{\Phi^2(G)} \log \left(\frac{1440}{s(G)}\right)\log \left(\frac{14}{s(G)}\right)  + 3 \log \left(\frac{14}{s(G)}\right) + 3
	\end{equation*}
\end{proposition}
\begin{proof}
	Fix $a = \frac{1}{18}$, and let $\tau_0 = \frac{2}{\Phi^2(G)} \log \left(\frac{80}{a s(G)}\right)$. Then,
	\begin{equation*}
	\left(1 - \frac{\Phi^2(G)}{2}\right)^{\tau_0} \leq \exp(- \tau_0\Phi^2(G)/2) \leq \frac{as(G)}{80}
	\end{equation*}
	and so by Proposition \ref{prop: tv_mixing_time},
	\begin{equation*}
	\norm{q_v^{(\tau_0 + 3)} - \pi}_{TV} \leq \max \set{\frac{1}{20}, \frac{1}{8} + \frac{1}{20} + \frac{1}{20}} + \left(\frac{2}{as(G)}\right) \frac{as(G)}{80} \leq \frac{1}{4}.
	\end{equation*}

	It is a well known fact \cite{montenegro2002} that if $\norm{q_v^{(t)} - \pi}_{TV} \leq \frac{1}{4}$, then for any $0 < \epsilon < 1$, $\norm{q_v^{(t\log_2(1/\epsilon))} - \pi}_{TV} \leq \epsilon$. Therefore, letting $\tau_1 = (\tau_0 + 3) \log(\frac{14}{s(G)})$, 
	\begin{equation*}
	\norm{q_v^{(\tau_1)} - \pi}_{TV} \leq \frac{1}{14 s(G)}
	\end{equation*}
	and so by Lemma \ref{lem: tv_to_uniform_distance}, $d_{\textrm{unif}}(q_v^{(\tau_1 + 3)}, \pi) \leq \frac{1}{4}$. 
\end{proof}

@@

\paragraph{Proof of Proposition \ref{prop: tv_mixing_time}.}
For any $S \subseteq V$ and any integer $t \geq 0$,
\begin{equation*}
\max \set{h_v^{(t)}(\pi(S)), h_v^{(t)}(1 - \pi(S))} \geq \abs{q_v^{(t)}(S) - \pi(S)}
\end{equation*}
and taking the maximum over all $S \subseteq V$ on both sides, we have
\begin{equation*}
\max_{0 \leq x \leq 1} h_v^{(t)}(x) \geq \norm{q_v^{(t)} - \pi}_{TV}.
\end{equation*}

@@

We will also need an upper bound on the distance between transition probability densities $\widetilde{P}_{\nu,r}(u,\cdot)$ and $\widetilde{P}_{\nu,r}(v,\cdot)$ for $u,v \in \Csig$ close together. 
\begin{lemma}[One-step distributions]
	\label{lem: one_step_distributions}
	Let $u,v \in \Csig$ be such that 
	\begin{equation*}
	\norm{u - v} \leq \frac{r t}{2\sqrt{d}}
	\end{equation*}
	for some $0 < t < 1/8$, and further assume there exists $\ell > 0$ such that $\ell_{\nu,r}(u), \ell_{\nu,r}(v) \geq \ell \nu_d r^d$. Then,
	\begin{equation*}
	\norm{\widetilde{P}_{\nu,r}(u; \cdot) - \widetilde{P}_{\nu,r}(v; \cdot)}_{TV} \leq 1 + \frac{3 \sqrt{3} t}{4\sqrt{2\pi}} - \ell
	\end{equation*}
	where $\norm{P - Q}_{TV}$ is the total variation distance between probabilities $P$ and $Q$. 
\end{lemma}

@@

\begin{proof}[Proof of Lemma \ref{lem: one_step_distributions}]
	Let $\Sset_1 \cup \Sset_2 = \Csig$ be an arbitrary partition of $\Csig$. We will show that 
	\begin{equation*}
	\widetilde{P}_{\nu,r}(u; \Sset_1) - \widetilde{P}_{\nu,r}(v; \Sset_1) \leq 1 + \frac{3 \sqrt{3} t}{4\sqrt{2\pi}} - \ell.
	\end{equation*}
	Since this will hold for arbitrary $\Sset_1 \in \mathfrak{B}(\Csig)$, it will hold for the infimum over all such $\Sset_1$ as well, and therefore the same lower bound will hold for $\norm{\widetilde{P}_{\nu,r}(u; \cdot) - \widetilde{P}_{\nu,r}(v; \cdot)}_{TV}$.

@@

So we can assume $\pi_{\nu,r}(\Sset_1') \geq \pi_{\nu,r}(\Sset_1)/2$, and likewise for $\Sset_2$. Now, for every $u \in \Sset_1', v \in \Sset_2'$, we have that
\begin{equation*}
\norm{\widetilde{P}_{\nu,r}(u;\cdot) - \widetilde{P}_{\nu,r}(v;\cdot)}_{TV} \geq 1 - \widetilde{P}_{\nu,r}(u;\Sset_1) - \widetilde{P}_{\nu,r}(v;\Sset_2) > 1 - \frac{\ell}{2}.
\end{equation*}

@@

\subsection{Proof of Theorem \ref{thm: misclassification_rate}}
\label{sec: proof_of_misclassification_rate}

In this section, we prove both Theorem \ref{thm: misclassification_rate} and the relevant portions of Corollary \ref{cor: appr}.

@@

\begin{lemma}
	\label{lem: zhu}
	Let $\varepsilon \geq 0$, $v \in S^g$, $\alpha \leq \frac{1}{9\tau_{\infty}(G[S])}$, and $\vol_0 \in [1 - c,1 + c] \vol(S;G)$ for $0 < c < 1$. Consider taking sweep cuts of the approximate \pprspace vector $\pbf^{(\varepsilon)} := \pbf^{(\varepsilon)}(v,\alpha; G)$ of the form
	\begin{equation*}
	S_{\beta} := \left\{u \in V: \pbf^{(\varepsilon)} \geq \beta \frac{\deg(u; G)}{\vol_0} \right\}
	\end{equation*}

	For any $\beta < \frac{3}{5(1 + c)}$, and $0 \leq \varepsilon \leq \frac{1}{10} \vol(S;G)$ the following bounds hold:
	\begin{align}
	\vol(S_{\beta} \setminus S; G) & \leq \frac{2\Phi(A;G)}{\alpha \beta(1 - c)} \vol(S; G), \label{eqn: zhu_1}\\
	\vol(S \setminus S_{\beta}; G) & \leq \left(\frac{2\Phi(S;G)}{\alpha (\frac{3}{5} - \beta(1 + c))} + 8\Phi(S;G)\right)\vol(S; G).  \label{eqn: zhu_2}
	\end{align}
\end{lemma}

@@

The choice $\varepsilon = \frac{1}{20 \vol_0}$, along with Lemma \ref{lem: ball_bounds_in_probability} and the upper bound on $\vol_0$ given by \eqref{eqn: initialization}, together imply that $\varepsilon < \frac{1}{10}\vol_{n,r}(\Csig[\Xbf])$ with probability tending to one as $n \to \infty$. We may therefore apply Lemma \ref{lem: zhu} with respect to $\pbf^{(\varepsilon)}$. All other aspects of the proof of Theorem \ref{thm: misclassification_rate} may be applied exactly as stated.

\subsection{Proof of Theorem \ref{thm: consistent_recovery_of_density_clusters}}
\label{sec: proof_of_consistent_cluster_recovery}

>>>>

In this paper, we study the problem of clustering: splitting a given data set
into groups that satisfy some notion of within-group similarity and
between-group difference.  We focus on spectral clustering, a family of powerful
nonparametric clustering algorithms.  Generally speaking, a spectral algorithm
first constructs a geometric graph $G$, where vertices correspond to samples,
and edges correspond to proximities between samples. It then estimates a feature
embedding based on the Laplacian of $G$, and applies a simple clustering
technique (like k-means clustering) in the embedded feature space.

When applied to geometric graphs built from a large number of samples,
global spectral clustering methods can be computationally cumbersome and   
insensitive to the local geometry of the underlying distribution
\citep{leskovec2010,mahoney2012}.  This has led to increased interest in
local spectral clustering algorithms, which leverage locally-biased spectra
computed using random walks around some user-specified seed node.  A popular 
local clustering algorithm is Personalized PageRank (PPR), first introduced by 
\citet{haveliwala2003}, then further developed by
several others~\citep{spielman2011,spielman2014,andersen2006,mahoney2012,zhu2013}.

Local spectral clustering techniques have been practically very successful
\citep{leskovec2010,andersen2012,gleich2012,mahoney2012,wu2012}, leading 
many authors to develop supporting theory
\citep{spielman2013,andersen2009,gharan2012,zhu2013} that gives worst-case
guarantees on traditional graph-theoretic notions of cluster quality (such as
conductance).  In this paper, we adopt a more traditional statistical viewpoint,
and examine what the output of local clustering on a data set reveals about the
underlying density of the samples 
$f$.  In particular, we examine the ability of PPR to recover
\emph{density clusters} of $f$, defined as the connected components of
the upper level set $\{x \in \Rd : f(x) \geq \lambda\}$ for some $\lambda > 0$
(a central object of interest in the statistical clustering literature, dating
back to the work of \citet{hartigan1981}).

@@

\paragraph{Summary of results.} A summary of our results (and an outline for this
paper) is as follows.

@@

More broadly, density clustering and level set estimation is a well-studied
problem. \citet{polonik1995, rigollet2009} study density clustering under the 
symmetric set difference metric, \citet{tsybakov1997, singh2009} describe
minimax optimal level-set estimators under Hausdorff loss and
\citet{hartigan1981, chaudhuri2010, balakrishnan2013} consider consistent estimation of the
cluster tree.
%, to note but a few works. 
Our goal is not to improve on these
results, nor to offer a better algorithm for level set estimation; indeed, seen as
a density clustering algorithm, PPR has none of the optimality guarantees 
found in the aforementioned works. Instead, our motivation is to start with a 
widely-used local spectral method, PPR, and to better understand and
characterize the distinctions between those density clusters which are
well-conditioned for PPR, and those which are not.

@@

\begin{theorem}
  \label{thm: misclassification_rate}
  Fix $\lambda > 0$, let \smash{$\Cset \in \Cbb_f(\lambda)$} be a
  $\kappa$-well-conditioned density cluster (with respect to some $\theta$), and
  additionally assume that $f$ satisfies \ref{asmp: valid_region}. If Algorithm
  \ref{alg: ppr} is well-initialized, there exists a universal constant $c >
  0$ such that with probability tending to 1 as $n \to \infty$,  
  \begin{equation}
    \label{eqn: misclassification_rate_ub}
    \Delta(\Csig[\Xbf], \Cest) \leq c \kappa(\Cset)
    \frac{\Lambda_{\sigma}}{\lambda_{\min}}. 
  \end{equation}
\end{theorem}

@@

Neither \eqref{eqn: graph_symmetric_set_difference} nor \eqref{eqn:
  misclassification_rate_ub} imply consistent density cluster estimation in the 
sense of \eqref{eqn: consistent_density_cluster_recovery}. This notion of
consistency requires a uniform bound over $\pbf$: for all \smash{$\Cset'
  \in \Cbb_f(\lambda), \Cset' \neq \Cset$}, and each $u \in \Cset, w \in
\Cset'$,  
\begin{equation}
\label{eqn: ppr_gap}
\frac{p_{w}}{\Dbf_{ww}} \leq \frac{1}{40\vol_0} < \frac{1}{11\vol_0} \leq
\frac{p_u}{\Dbf_{uu}}, 
\end{equation}
so that any sweep cut $S_{\beta}$ for $\beta \vol_0 \in [1/40,1/11]$ (i.e., any
sweep cut considered by Algorithm \ref{alg: ppr}) will fulfill both conditions
laid out in \eqref{eqn: consistent_density_cluster_recovery}. In Theorem
\ref{thm: consistent_recovery_of_density_clusters}, we show that a sufficiently
small upper bound on $\kappa(\Cset)$ ensures that such a gap exists with probability
1 as $n \to \infty$, and hence guarantees \smash{$\Cest$} will be a consistent  
estimator. As was the case before, we wish to preclude arbitrarily low degree
vertices, this time for points $x \in \Cset'[\Xbf]$. 
\begin{enumerate}[label=(A\arabic*)]
  \setcounter{enumi}{6}
\item 
  \label{asmp: C'_bounded_density}
  \emph{Bounded density in other clusters:} Letting $\sigma,\lambda_{\sigma}$ be 
  as in \ref{asmp: bounded_density}, for each $\Cset' \in \Cbb_f(\lambda)$ and
  for all $x \in \Csig'$, $\lambda_{\sigma} \leq f(x)$. 
\end{enumerate}

@@

We generalize some notation from the main text. For an undirected graph $G = (V,E)$, the lazy random walk over $G$ is the Markov chain with transition probabilities given by $\Wbf := \frac{\Ibf{} + \Dbf^{-1} \Abf}{2}$ and stationary distribution $\pi$.  Denote the $m$-step probability distribution of this random walk originating from a particular $v \in V$ as $q^{(m)}: V \times V \to [0,1]$, $q^{(m)}(v,u) = e_v \Wbf^m e_u$. An important intermediate quantity used to bound the relative pointwise mixing time $\tau_{\infty}(G)$ is the total variation distance between the distributions $q_v^{(m)} := q^{(m)}(v,\cdot)$ and $\pi$,
\begin{equation*}
\norm{q_v^{(m)} - \pi}_{\mathrm{TV}} = \sum_{u \in V} \abs{q_v^{(m)}(u) - \pi(u)}
\end{equation*}
We recall the \emph{cut} and \emph{volume} functionals over a graph, and introduce the \emph{degree} functional as well. For $u \in V$, $S \subseteq V$, 
\begin{equation*}
\cut(S;G) = \sum_{u \in S} \sum_{v \in S^c} \1((u,v) \in E), \quad \deg(u;G) = \sum_{v \in V} \1((u,v) \in E), \quad \vol(S;G) = \sum_{u \in S} \deg(u;G)
\end{equation*}

@@

Proposition \ref{prop: tv_mixing_time} relates these geometric quantities to the TV distance between $q_v^{(t)}$ and stationary distribution $\pi$.
\sbcomment{Added $\max$ to the RHS.}
\begin{proposition}
	\label{prop: tv_mixing_time}
	For any $v \in V$, and any $0 < a \leq 1$,
	\begin{align*}
	\norm{q_v^{(t + 3)} - \pi}_{\mathrm{TV}} \leq \max\set{as(G), \frac{1}{8} + \frac{1}{20} + \frac{1}{2 \min_{u \in V} \deg(u;G)}} 
	 \\
	~~~~~~~~~~~~~~~~~~~~~ + \left(\frac{1}{1 - 2as(G)} +  \frac{1}{2 as(G)}\right) \left(1 - \frac{\Phi^2(G)}{2}\right)^t
	\end{align*}
\end{proposition}

@@

\begin{lemma}
	\label{lem: tv_to_uniform_distance}
	Let $\norm{q_v^{(t)} - \pi}_{\mathrm{TV}} \leq \frac{1}{14} \max \set{1, \frac{1}{s(G)}}$. Then,
	\begin{equation*}
	d_{\mathrm{unif}}(q_v^{(t + 3)},\pi) \leq \frac{1}{4}
	\end{equation*}
\end{lemma}
\begin{proof}
	Fix $u \in V$ and let $m \geq t + 1$ be arbitrary. The stationarity of $\pi$ gives
	\begin{align}
	\frac{\pi(u) - q_v^{m}(u)}{\pi(u)} & = \sum_{y \in V} \left(\pi(y) - q^{(m-1)}(v,y)\right) \left(\frac{q^{(1)}(y,u) - \pi(u)}{\pi(u)}\right) \nonumber \\
	& \overset{\text{(i)}}{=} \sum_{y \neq u} \left(\pi(y) - q^{(m-1)}(v,y)\right) \left(\frac{q^{(1)}(y,u) - \pi(u)}{\pi(u)}\right) + \frac{\pi(u) - q^{(m - 1)}(v,u)}{\pi(u)} \left(\frac{1}{2} - \pi(u)\right) \nonumber \\
	& \leq \sum_{y \neq u} \left(\pi(y) - q^{(m-1)}(v,y)\right) \left(\frac{q^{(1)}(y,u) - \pi(u)}{\pi(u)}\right) + \frac{\pi(u) - q^{(m - 1)}(v,u)}{2 \pi(u)} \label{eqn: tv_to_uniform_distance_1}
	\end{align}
	where $\text{(i)}$ follows from $q^{(1)}(u,u) = \frac{1}{2}$.

	Then 
	\begin{align}
	\sum_{y \neq u} \left(\pi(y) - q^{(m-1)}(v,y)\right) \left(\frac{q^{(1)}(y,u) - \pi(u)}{\pi(u)}\right) & \leq \norm{q_v^{(m-1)} - \pi}_{\mathrm{TV}} \max_{y \neq u} \abs{\frac{q^{(1)}(y,u) - \pi(u)}{\pi(u)}} \nonumber \\
	& \leq \norm{q_v^{(m-1)} - \pi}_{\mathrm{TV}} \max \set{1, \max_{y \neq u}\set{\frac{q^{(1)}(y,u)}{\pi(u)}}} \nonumber \\
	& \leq \norm{q_v^{(m-1)} - \pi}_{\mathrm{TV}} \max \set{1, \frac{1}{s(G)}} \label{eqn: tv_to_uniform_distance_3}
	\end{align}
	since for $y \neq u$, $q^{(1)}(y,u) \leq 1/\left(2 \min_{u \in V} \deg(u; G)\right)$. As $m - 1 \geq t$, it is well known \cite{lovasz1990} that the laziness of the random walk guarantees $\norm{q_v^{(m - 1)} - \pi}_{\mathrm{TV}} \leq \norm{q_v^{(t)} - \pi}_{\mathrm{TV}}$, and therefore by \eqref{eqn: tv_to_uniform_distance_3} and hypothesis,
	\begin{equation*}
	\sum_{y \neq u} \left(\pi(y) - q^{(m-1)}(v,y)\right) \left(\frac{q^{(1)}(y,u) - \pi(u)}{\pi(u)}\right) \leq \frac{1}{14}.
	\end{equation*}

@@

\begin{proposition}
	\label{prop: pointwise_mixing_time}
	Assume $\min_{u \in v} \deg(u; G) \geq 10$. Then,
	\begin{equation*}
	\tau_{\infty}(G) \leq \frac{2}{\Phi^2(G)} \log \left(\frac{1440}{s(G)}\right)\log \left(\frac{14}{s(G)}\right)  + 3 \log \left(\frac{14}{s(G)}\right) + 3
	\end{equation*}
\end{proposition}
\begin{proof}
	Fix $a = \frac{1}{18}$, and let $\tau_0 = \frac{2}{\Phi^2(G)} \log \left(\frac{80}{a s(G)}\right)$. Then,
	\begin{equation*}
	\left(1 - \frac{\Phi^2(G)}{2}\right)^{\tau_0} \leq \exp(- \tau_0\Phi^2(G)/2) \leq \frac{as(G)}{80}
	\end{equation*}
	and so by Proposition \ref{prop: tv_mixing_time},
	\begin{equation*}
	\norm{q_v^{(\tau_0 + 3)} - \pi}_{\mathrm{TV}} \leq \max \set{\frac{1}{20}, \frac{1}{8} + \frac{1}{20} + \frac{1}{20}} + \left(\frac{2}{as(G)}\right) \frac{as(G)}{80} \leq \frac{1}{4}.
	\end{equation*}

	It is a well known fact \cite{montenegro2002} that if $\norm{q_v^{(t)} - \pi}_{\mathrm{TV}} \leq \frac{1}{4}$, then for any $0 < \epsilon < 1$, $\norm{q_v^{(t\log_2(1/\epsilon))} - \pi}_{\mathrm{TV}} \leq \epsilon$. Therefore, letting $\tau_1 = (\tau_0 + 3) \log(\frac{14}{s(G)})$, 
	\begin{equation*}
	\norm{q_v^{(\tau_1)} - \pi}_{\mathrm{TV}} \leq \frac{1}{14 s(G)}
	\end{equation*}
	and so by Lemma \ref{lem: tv_to_uniform_distance}, $d_{\textrm{unif}}(q_v^{(\tau_1 + 3)}, \pi) \leq \frac{1}{4}$. 
\end{proof}

@@

\paragraph{Proof of Proposition \ref{prop: tv_mixing_time}.}
For any $S \subseteq V$ and any integer $t \geq 0$,
\begin{equation*}
\max \set{h_v^{(t)}(\pi(S)), h_v^{(t)}(1 - \pi(S))} \geq \abs{q_v^{(t)}(S) - \pi(S)}
\end{equation*}
and taking the maximum over all $S \subseteq V$ on both sides, we have
\begin{equation*}
\max_{0 \leq x \leq 1} h_v^{(t)}(x) \geq \norm{q_v^{(t)} - \pi}_{\mathrm{TV}}.
\end{equation*}

@@

We will also need an upper bound on the distance between transition probability densities $\widetilde{P}_{\nu,r}(u,\cdot)$ and $\widetilde{P}_{\nu,r}(v,\cdot)$ for $u,v \in \Csig$ close together. 
\begin{lemma}[One-step distributions]
	\label{lem: one_step_distributions}
	Let $u,v \in \Csig$ be such that 
	\begin{equation*}
	\norm{u - v} \leq \frac{r t}{2\sqrt{d}}
	\end{equation*}
	for some $0 < t < 1/8$, and further assume there exists $\ell > 0$ such that $\ell_{\nu,r}(u), \ell_{\nu,r}(v) \geq \ell \nu_d r^d$. Then,
	\begin{equation*}
	\norm{\widetilde{P}_{\nu,r}(u; \cdot) - \widetilde{P}_{\nu,r}(v; \cdot)}_{\mathrm{TV}} \leq 1 + \frac{3 \sqrt{3} t}{4\sqrt{2\pi}} - \ell
	\end{equation*}
	where $\norm{P - Q}_{\mathrm{TV}}$ is the total variation distance between probabilities $P$ and $Q$. 
\end{lemma}

@@

\begin{proof}[Proof of Lemma \ref{lem: one_step_distributions}]
	Let $\Sset_1 \cup \Sset_2 = \Csig$ be an arbitrary partition of $\Csig$. We will show that 
	\begin{equation*}
	\widetilde{P}_{\nu,r}(u; \Sset_1) - \widetilde{P}_{\nu,r}(v; \Sset_1) \leq 1 + \frac{3 \sqrt{3} t}{4\sqrt{2\pi}} - \ell.
	\end{equation*}
	Since this will hold for arbitrary $\Sset_1 \in \mathfrak{B}(\Csig)$, it will hold for the infimum over all such $\Sset_1$ as well, and therefore the same lower bound will hold for $\norm{\widetilde{P}_{\nu,r}(u; \cdot) - \widetilde{P}_{\nu,r}(v; \cdot)}_{\mathrm{TV}}$.

@@

So we can assume $\pi_{\nu,r}(\Sset_1') \geq \pi_{\nu,r}(\Sset_1)/2$, and likewise for $\Sset_2$. Now, for every $u \in \Sset_1', v \in \Sset_2'$, we have that
\begin{equation*}
\norm{\widetilde{P}_{\nu,r}(u;\cdot) - \widetilde{P}_{\nu,r}(v;\cdot)}_{\mathrm{TV}} \geq 1 - \widetilde{P}_{\nu,r}(u;\Sset_1) - \widetilde{P}_{\nu,r}(v;\Sset_2) > 1 - \frac{\ell}{2}.
\end{equation*}

@@

\section{Proof of Theorem \ref{thm: misclassification_rate}}
\label{sec: proof_of_misclassification_rate}
In this section, we prove both Theorem \ref{thm: misclassification_rate} and the relevant portions of Corollary \ref{cor: appr}.

@@

\begin{lemma}
	\label{lem: zhu}
	Let $\varepsilon \geq 0$, $v \in S^g$, $\alpha \leq \frac{1}{9\tau_{\infty}(G[S])}$, and $\vol_0 \in [1 - c,1 + c] \vol(S;G)$ for $0 < c < 1$. Consider taking sweep cuts of the approximate \pprspace vector $\pbf^{(\varepsilon)} := \pbf^{(\varepsilon)}(v,\alpha; G)$ of the form
	\begin{equation*}
	S_{\beta} := \left\{u \in V: \pbf^{(\varepsilon)} \geq \beta \frac{\deg(u; G)}{\vol_0} \right\}
	\end{equation*}
	For any $\beta < \frac{3}{5(1 + c)}$, and $0 \leq \varepsilon \leq \frac{1}{10} \vol(S;G)$ the following bounds hold:
	\begin{align}
	\vol(S_{\beta} \setminus S; G) & \leq \frac{2\Phi(A;G)}{\alpha \beta(1 - c)} \vol(S; G), \label{eqn: zhu_1}\\
	\vol(S \setminus S_{\beta}; G) & \leq \left(\frac{2\Phi(S;G)}{\alpha (\frac{3}{5} - \beta(1 + c))} + 8\Phi(S;G)\right)\vol(S; G).  \label{eqn: zhu_2}
	\end{align}
\end{lemma}

@@

The choice $\varepsilon = \frac{1}{20 \vol_0}$, along with Lemma~\ref{lem: ball_bounds_in_probability} and the upper bound on $\vol_0$ given by \eqref{eqn: initialization}, together imply that $\varepsilon < \frac{1}{10}\vol_{n,r}(\Csig[\Xbf])$ with probability tending to one as $n \to \infty$. We may therefore apply Lemma \ref{lem: zhu} with respect to $\pbf^{(\varepsilon)}$. All other aspects of the proof of Theorem \ref{thm: misclassification_rate} may be applied exactly as stated.

\section{Proof of Theorem \ref{thm: consistent_recovery_of_density_clusters}}
\label{sec: proof_of_consistent_cluster_recovery}
